---
title: "kestrel data"
author: "Trevor Caughlin"
date: '2022-04-14'
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fitdistrplus)
library("HyperbolicDist")
```

## Dispersal

In this presentation, we will be modeling dispersal of American Kestrels using data from McCaslin 2020.

## Data details

- Banding and encounter data from the US Geological Survey's Bird Banding Laboratory
- All kestrels banded in the United States or Canada from 1961 to 2015
- n=329,987 kestrels banded,5,329 (1.6%) of those birds were subsequently reported. 
- define natal dispersers as ‘local’ (nestling) or ‘hatch year’ birds banded during the breeding season (1 April–15 August) and encountered during the breeding season 1 year later.

## Load the data
```{r cars, echo = TRUE}
kestrels=read.csv("mccaslin_kestrel_final_dataset_archive.csv")
hist(kestrels$dist_km)
```

What do these data tell us about natal dispersal for kestrels?

## Fit gamma regression
To quantify effects of covariates, we will use a gamma regression to model distance.
Note that the original paper used a cut off of 30 km to define long vs. short dispersal.

```{r gamma}
ldd_kestrels<-kestrels[which(kestrels$dist_km>30),]
hist(ldd_kestrels$dist_km)
#extB - percentage of agricultural land cover in the 4 sqkm area surrounding the encounter lcoation
gamma_model<-glm(dist_km~extE,family = Gamma(link = "log"),data=ldd_kestrels)
summary(gamma_model)
```

What are some of the advantages and disadvantages of modeling with a gamma distribution?

**In class activity: fit a gamma model to one or more of the covariates. What do your results show?**

## How much better do the fat-tailed distributions fit, relative to the gamma?

```{r distribution}

dist_nonzero<-kestrels$dist_km[which(kestrels$dist_km>0)]
#removing zero values to enable fit of gamma

gamma_fit<-fitdist(dist_nonzero,distr="gamma")

#visually explore fit:
curve(dgamma(x,shape=gamma_fit[1]$estimate[1],rate=gamma_fit$estimate[2]),from=0,to=1000,xlab="Dispersal distance (km)",ylab="Probability density")


log_sech_fit<-hyperbFit(dist_nonzero)

curve(dhyperb(x,Theta=log_sech_fit$Theta),from=0,to=1000,add=T,col="red")

#calculate fit using AIC

AIC_log_sech_fit<-2*4-2*log_sech_fit$maxLik
AIC_gamma<-gamma_fit$aic

```

**Assignment**: using the half-cauchy distribution in the extraDistr package [https://www.rdocumentation.org/packages/extraDistr/versions/1.9.1/topics/HalfCauchy] fit a half-cauchy distribution to the kestral natal dispersal data.

1) calculate AIC and compare with the log-sech and gamma distribution. Which distribution provides the best fit to the data?

2) Add a curved line for the half-cauchy to the plot. How does the visual fit of the data compare?

